`mrpkit`: A reproducible, opinionated, workflow for multilevel regression with post-stratification in R

### 2021-12-07 paper/paper.Rmd
Add the US election example to the paper

<!-- Probs get rid of: -->
<!-- The `SurveyMap` object has `new`, `add`, `delete`, `replace`, `mapping`, `tabulate`, and `fit` methods. -->

@@

Once we have the aggregated estimates, we can visualize them using the `plot` method. This method takes the aggregated data frame and additional_stats option (the default is wtd and raw estimates) as its arguments. It generates a violin plot of the aggregated estimate by the level of a certain variable. When additional statistics are specified (options are raw, wtd and mrp; more than one can be specified as a vector), these are added as points with error bars representing the 95% confidence interval.  
In this example, we display the estimated plot for `age` and `state` and a density plot for the population estimates.

@@

The next plot displays the probability of voting for the BP by the state without using additional statistics. We learn that the probability of the BP winning the election is high since most people in almost all the states (around 75%) vote for that party. State B is the state with the lowest probability of vote for the BP. The distribution still ranged more than 50%. Note that this violin plot is generated using `ggplot2` (Wickham, 2016). Hence, it could be modified the same way `ggplot2`'s plot is modified. For example, here, we can add the title and change the theme of the plot.

@@

@@

>>>>

@@

Once we have the aggregated estimates, we can visualize them using the `plot` method. This method takes the aggregated data frame and additional_stats option (the default is wtd and raw estimates) as its arguments. It generates a violin plot of the aggregated estimate by the level of a certain variable. When additional statistics are specified (options are raw, wtd and mrp; more than one can be specified as a vector), these are added as points with error bars representing the 95 per cent confidence interval.  
In this example, we display the estimated plot for 'age' and 'state' and a density plot for the population estimates.

@@

The next plot displays the probability of voting for the BP by the state without using additional statistics. We learn that the probability of the BP winning the election is high since most people in almost all the states (around 75%) vote for that party. State B is the state with the lowest probability of vote for the BP. The distribution still ranged more than 50 per cent. Note that this violin plot is generated using `ggplot2` [@citeggplot]. Hence, it could be modified the same way `ggplot2`'s plot is modified. For example, here, we can add the title and change the theme of the plot.

@@

There are a variety of technical notes related to this package and the architectural decisions that were made.

Section \@ref(implementation) discusses some of the implementation issues and technical notes related to the decisions that were made

@@

In this example we will forecast the 2020 US presidential election using survey data from the Democracy Fund Voter Study Group and use the Integrated Public Use Microdata Series (IPUMS), to access the 2018 American Community Survey (ACS) to use as post-stratification data. We are interested in considering the following variables: state, age-group, and gender.

The first step is to gather the datasets. For the survey data we are interested in the Nationscape dateset, which can be requested from the Voter Study Group. Access is usually granted within hours and is completely free. We are going to use the survey from 25 June 2020.

```{r}
# This code based on that from Alen Mitrovski, Xiaoyan Yang, and Matthew Wankiewicz.
library(haven)
library(tidyverse)

raw_nationscape_data <- 
  read_dta(here::here("paper/ns20200625.dta"))

# The Stata format separates labels so reunite those
raw_nationscape_data <- 
  labelled::to_factor(raw_nationscape_data)

# Just keep relevant variables
nationscape_data <- 
  raw_nationscape_data %>% 
  select(vote_2020,
         gender,
         state,
         age)

# For simplicity, remove anyone undecided or planning to vote for someone other than Biden/Trump and make vote a binary variable: 1 for Biden, 0 for Trump.
nationscape_data <- 
  nationscape_data %>% 
  filter(vote_2020 == "Joe Biden" | vote_2020 == "Donald Trump") %>% 
  mutate(vote_biden = if_else(vote_2020 == "Joe Biden", 1, 0)) %>% 
  select(-vote_2020)

# Create the dependent variables by grouping the existing variables
nationscape_data <-
  nationscape_data %>%
  mutate(
    age_group = case_when( # case_when works in order and exits when there's a match
      age <= 29 ~ 'age_18-29',
      age <= 44 ~ 'age_30-44',
      age <= 59 ~ 'age_45-59',
      age >= 60 ~ 'age_60_or_more',
      TRUE ~ 'Trouble'
      )
#     gender = case_when(
#       gender == "Female" ~ 'female',
#       gender == "Male" ~ 'male',
#       TRUE ~ 'Trouble'
#       ),
#     education_level = case_when(
#       education == "3rd Grade or less" ~ "High school or less",
#       education == "Middle School - Grades 4 - 8" ~ "High school or less",
#       education == "Completed some high school" ~ "High school or less",
#       education == "High school graduate" ~ "High school or less",
#       education == "Other post high school vocational training" ~ "Some post secondary",
#       education == "Completed some college, but no degree" ~ "Some post secondary",
#       education == "Associate Degree" ~ "Post secondary or higher",
#       education == "College Degree (such as B.A., B.S.)" ~ "Post secondary or higher",
#       education == "Completed some graduate, but no degree" ~ "Post secondary or higher",
#       education == "Masters degree" ~ "Graduate degree",
#       education == "Doctorate degree" ~ "Graduate degree",
#       TRUE ~ 'Trouble'
#       )
    ) 
#   select(-education, -age)

# Tidy the class
nationscape_data <-
  nationscape_data %>%
  mutate(across(c(gender, state, age_group, vote_biden), as_factor))

nationscape_data %>% 
  head()
```

```{r}
# This code is very directly from Alen Mitrovski, Xiaoyan Yang, and Matthew Wankiewicz.
# Format state names so the whole state name is written out, to match IPUMS data
# states_names_and_abbrevs <- 
#   tibble(stateicp = state.name, state = state.abb)
# 
# nationscape_data <-
#   nationscape_data %>%
#   left_join(states_names_and_abbrevs)
# 
# rm(states_names_and_abbrevs)
# 
# # Make lowercase to match IPUMS data
# nationscape_data <- 
#   nationscape_data %>% 
#   mutate(stateicp = tolower(stateicp))
# 
# nationscape_data$stateicp %>% unique()
# 
# # Replace NAs with DC
# nationscape_data$stateicp <- 
#   replace_na(nationscape_data$stateicp, "district of columbia")
# 
# # Tidy the class
# nationscape_data <- 
#   nationscape_data %>% 
#   mutate(across(c(gender, stateicp, education_level, age_group), as_factor))
# 
# # Save data
# write_csv(nationscape_data, "outputs/data/polling_data.csv")

We also need a post-stratification dataset. In this case we can use the American Community Survey (ACS) accessed through IPUMS.

```{r}
# Again, closely following code from Alen Mitrovski, Xiaoyan Yang, and Matthew Wankiewicz.
library(haven)
library(tidyverse)

raw_poststrat_data <- 
  read_dta(here::here("paper/usa_00009.dta"))

# The Stata format separates labels so reunite those
raw_poststrat_data <- 
  labelled::to_factor(raw_poststrat_data)
head(raw_poststrat_data)

raw_poststrat_data$age <- as.numeric(raw_poststrat_data$age)
raw_poststrat_data$stateicp <- as.character(raw_poststrat_data$stateicp)

poststrat_data <- 
  raw_poststrat_data %>% 
  # filter(inctot < 9999999) %>% 
  filter(age >= 18) %>% 
#   mutate(gender = sex) %>% 
  mutate(
    age_group = case_when( # case_when works in order and exits when there's a match
      age <= 29 ~ 'age_18-29',
      age <= 44 ~ 'age_30-44',
      age <= 59 ~ 'age_45-59',
      age >= 60 ~ 'age_60_or_more',
      TRUE ~ 'Trouble'
      )
#     education_level = case_when(
#       educd == "nursery school, preschool" ~ "High school or less",
#       educd == "kindergarten" ~ "High school or less",
#       educd == "grade 1" ~ "High school or less",
#       educd == "grade 2" ~ "High school or less",
#       educd == "grade 3" ~ "High school or less",
#       educd == "grade 4" ~ "High school or less",
#       educd == "grade 5" ~ "High school or less",
#       educd == "grade 6" ~ "High school or less",
#       educd == "grade 7" ~ "High school or less",
#       educd == "grade 8" ~ "High school or less",
#       educd == "grade 9" ~ "High school or less",
#       educd == "grade 10" ~ "High school or less",
#       educd == "grade 11" ~ "High school or less",
#       educd == "12th grade, no diploma" ~ "High school or less",
#       educd == "regular high school diploma" ~ "High school or less",
#       educd == "ged or alternative credential" ~ "High school or less",
#       educd == "some college, but less than 1 year" ~ "Some post secondary",
#       educd == "1 or more years of college credit, no degree" ~ "Some post secondary",
#       educd == "associate's degree, type not specified" ~ "Post secondary or higher",
#       educd == "bachelor's degree" ~ "Post secondary or higher",
#       educd == "master's degree" ~ "Graduate degree",
#       educd == "professional degree beyond a bachelor's degree" ~ "Graduate degree",
#       educd == "doctoral degree" ~ "Graduate degree",
#       educd == "no schooling completed" ~ "High school or less",
#       TRUE ~ 'Trouble'
#       )
    ) %>% 
  filter(!stateicp %in% c("puerto rico", "state groupings (1980 urban/rural sample)", "military/mil. reservations", "state not identified"))

# Just keep relevant variables
poststrat_data <- 
  poststrat_data %>% 
  select(sex,
         age_group,
         stateicp)

# Tidy the class
poststrat_data <-
  poststrat_data %>%
  mutate(across(c(sex, stateicp, age_group), as_factor))

# 
# # Save data
# write_csv(poststrat_data, "outputs/data/us_poststrat.csv")
# 
# poststrat_data %>% 
#   head()

Now that we have the data, we need to create a `SurveyData` object for the survey dataset.

```{r survey-object}
library(mrpkit)
# Create a SurveyData object for the survey dataset 
us_presidential_election <- 
  SurveyData$new(
    data = nationscape_data,
    questions = list(
      age_group = "Please identify your age",
      gender = "Please select your gender",
      vote_biden = "If today is election day, will you vote for Biden?",
      state = "Which state do you live in?"
      ),
    responses = list(
      age_group = levels(nationscape_data$age_group),
      gender = levels(nationscape_data$gender),
      vote_biden = levels(nationscape_data$vote_biden),
      state = levels(nationscape_data$state)
      )
    )

# Display it
us_presidential_election$print()
```

And we now need the same for the post-stratification dataset.

```{r survey-object}
# Create a SurveyData object for the survey dataset 
us_post_strat <- 
  SurveyData$new(
    data = poststrat_data,
    questions = list(
      age_group = "Please identify your age",
      sex = "Please select your sex",
      stateicp = "Which state do you live in?"
      ),
    responses = list(
      age_group = levels(poststrat_data$age_group),
      sex = levels(poststrat_data$sex),
      stateicp = levels(poststrat_data$stateicp)
      )
    )

# Display it
us_post_strat$print()
```

Now we need to start to bring the datasets together.

```{r questions}
# Create QuestionMap$object for the question related to age
q_age <- QuestionMap$new(
  name = "age",
  col_names = c("age_group", "age_group"),
  values_map = list(
    "age_18-29" = "age_18-29",
    "age_30-44" = "age_30-44",
    "age_45-59" = "age_45-59",
    "age_60_or_more" = "age_60_or_more"
  )
)
# 
# # Create QuestionMap$object for the question related to party preference
# q_party_pref <- QuestionMap$new(
#   name = "party_pref",
#   col_names = c("vote_for", "vote_pref"),
#   values_map = list(
#     "Box Party" = "BP",
#     "BP" = "BP",
#     "Circle Party" = "CP",
#     "CP" = "CP"
#   )
# )

# Create QuestionMap$object for the question related to gender
q_gender <- QuestionMap$new(
  name = "gender",
  col_names = c("gender", "sex"),
  values_map = data.frame(
    "Female" = "female",
    "Male" = "male"
  )
)

# Create QuestionMap$object for the question related to state
q_state <- QuestionMap$new(
  name = "state",
  col_names = c("state", "stateicp"),
  values_map = list(
  "AL" = "alabama",
  "AK" = "alaska",
  "AZ" = "arizona",
  "AR" = "arkansas",
  "CA" = "california",
  "CO" = "colorado",
  "CT" = "connecticut",
  "DE" = "delaware",
  "DC" = "district of columbia",
  "FL" = "florida",
  "GA" = "georgia",
  "HI" = "hawaii",
  "ID" = "idaho",
  "IL" = "illinois",
  "IN" = "indiana",
  "IA" = "iowa",
  "KS" = "kansas",
  "KY" = "kentucky",
  "LA" = "louisiana",
  "ME" = "maine",
  "MD" = "maryland",
  "MA" = "massachusetts",
  "MI" = "michigan",
  "MN" = "minnesota",
  "MS" = "mississippi",
  "MO" = "missouri",
  "MT" = "montana",
  "NE" = "nebraska",
  "NV" = "nevada",
  "NH" = "new hampshire",
  "NJ" = "new jersey",
  "NM" = "new mexico",
  "NY" = "new york",
  "NC" = "north carolina",
  "ND" = "north dakota",
  "OH" = "ohio",
  "OK" = "oklahoma",
  "OR" = "oregon",
  "PA" = "pennsylvania",
  "RI" = "rhode island",
  "SC" = "south carolina",
  "SD" = "south dakota",
  "TN" = "tennessee",
  "TX" = "texas",
  "UT" = "utah",
  "VT" = "vermont",
  "VA" = "virginia",
  "WA" = "washington",
  "WV" = "west virginia",
  "WI" = "wisconsin",
  "WY" = "wyoming"
  )
)
```

```{r}
q_age
q_gender
q_state
```

We can now bring them together.

```{r map-object}
# Create a new SurveyMap object
us_map <- SurveyMap$new(sample = us_presidential_election, 
                        population = us_post_strat,
                        q_state, q_age, q_gender
                        )

print(us_map)
```

We now finish the process of bringing the datasets together.

```{r mapping}
us_map$mapping()
```

```{r tabulate}
us_map$tabulate()
```

We can now model the dataset.

```{r  fitting}
# Example of using rstanarm::stan_glmer
fit1 <- us_map$fit(
  fun = rstanarm::stan_glmer,
  formula = vote_biden ~ (1 | age) + (1 | gender) + (1 | state),
  family = "binomial",
  refresh = 100,
  cores = 2)
```

```{r predict}
# predict the probability of voting for Biden using the fit1 model
poststrat_est_fit1 <- fit1$population_predict()
```

```{r}
# aggregate the predicted value by state
state_estimation <- fit1$aggregate(poststrat_est_fit1, by = "state")

```{r}

# generate the population estimate
popn_estimation <- fit1$aggregate(poststrat_est_fit1)
```

```{r}
library(ggplot2)
plot_state <- fit1$plot(state_estimation, additional_stats = "none") +
  ggtitle("Probability of voting the BP by state") +
  theme_bw()
plot_state
```
